%//////////////////////////////////////////////////////////////////////////////

\documentclass[fontsize=10pt,paper=A4,pagesize,DIV=15]{scrartcl}

\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage[british]{babel}

%\usepackage{fixltx2e}
\usepackage{ellipsis}
\usepackage{ragged2e}
\usepackage[final]{microtype}

\usepackage{palatino}

\usepackage{overcite}
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{perpage}
\usepackage{rotating}
\usepackage{svg}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage[colorlinks=true,
            urlcolor=blue,
            pdftex,
            pdfsubject  = {},
            pdfauthor   = {Oliver Kowalke},
            pdfkeywords = {C++,callcc,call/cc,context,continuation,coroutine,execution,fiber,P0099,P0534},
            pdftitle    = {call/cc: A low-level API for stackful context switching}]{hyperref}

%//////////////////////////////////////////////////////////////////////////////

\input{commands}

%//////////////////////////////////////////////////////////////////////////////

\begin{document}
\small
\begin{tabbing}
    Document number: \= P0534R1\\
    Date:            \> 2017-06-17\\
    Reply-to:        \> Oliver Kowalke (oliver.kowalke@gmail.com)\\
    Audience:        \> SG1/LEWG\\
\end{tabbing}

\section*{call/cc (call-with-current-continuation): A low-level API for stackful context switching}

%//////////////////////////////////////////////////////////////////////////////

\tableofcontents

%//////////////////////////////////////////////////////////////////////////////

\abschnitt{Abstract}
This document proposes a C++ equivalent to the well-known concept
\bfs{call-with-current-continuation} (abbreviated \bfs{call/cc}). This
facility permits a program written in portable C++ to subdivide processing into
distinct \bfs{contexts:} units smaller than a thread.\\

Within this proposal, the unadorned term ``thread'' means a \cpp{std::thread}
(or \bfs{kernel thread}).
When the Standard's more general term ``thread of execution'' is intended, it
is spelled out in full.\\

With \cc, processing in a given thread may be further subdivided into multiple
contexts. Each such context qualifies as a ``thread of execution'' according
to the definition in the Standard. However, within a given thread, control is
cooperatively passed from one context to another.\\

This has a couple of important implications:

\begin{itemize}
\item In each thread in a process, exactly one context is running at any given
  time. All others are \bfs{suspended.}
\item The running context on a thread continues running until it explicitly
  \bfs{resumes} some other context. The act of resuming another context
  suspends the previously-running context. This transfer of control, in which
  one context suspends and another resumes, is \bfs{context-switching.}
\item There are no data races between contexts running on the same thread.
\end{itemize}

The kind of context-switching presented in this proposal is called
\bfs{stackful} because each context requires some implementation of the C++
stack. C++ code running on a particular context may transparently call
ordinary C++ functions. In contrast to the \bfs{\coawait}
facility (proposed separately\cite{N4649}), this permits encapsulation. A
function that suspends (by resuming some other context) needs no special
signature. Its caller need not be aware that it might suspend. It need not
call that function in any special way.\\

This supports use cases that cannot be addressed with \coawait alone.\\

Also in contrast to the \coawait facility, this proposal requires no
changes to the core C++ language. \cc is presented as a library facility,
albeit a library that cannot be implemented in portable C++. This is why it is
desirable to incorporate it into the International Standard.\\

Consider the following bullets from P0559R0:\cite{P0559R0}

\begin{itemize}
\item Avoid 'compiler magic' when possible
\item Prefer library solutions over language changes if feasible
\end{itemize}

The proposed \cc facility is intended to be foundational. While of course
application coders are free to use the \cc API, its real promise is in
supporting higher-level abstractions.\\

This proposal describes the basic \cc facility, presents some illustrative use
cases and explains why the API is set at its present level.

\abschnitt{Why should \cc be standardized?}

The \cc facility cannot itself be implemented in portable C++. The present
implementation,\cite{bcontext} maintained by a single author, supports a
small set of current platforms available to that author. Should \cc be
integrated into the Standard, it will become universally available.\\

Compiler awareness of this facility could enable certain optimizations as
well:

\begin{itemize}
  \item The compiler might be able to analyze the code to be launched on a new
    \cc context and determine an optimal stack size for that context.
  \item The compiler might be able to determine that not all registers need be
    preserved across a context switch.
  \item For certain use cases, the compiler might be able to optimize away
    context-switching altogether. Promising work has been done in this area
    for the \coawait facility.\cite{N4649}
\end{itemize}

\abschnitt{Revision History}
This document supersedes P0534R0.\cite{P0534R0}\\
\newline
Changes since P0534R0:

\begin{itemize}
    \item API modified
    \begin{itemize}
        \item \cpp{operator()} renamed to \resume
        \item \cpp{operator(invoke\_ontop\_arg\_t)} renamed to \resumewith
        \item \dataavail, \getdata and \anythread are now member functions
          instead of free functions
        \item stack unwinding now explicitly requires specific \cpp{unwind\_exception} 
              exception
    \end{itemize}
    \item \cc compared with \uc and \lj
    \item why \cc is a low-level implementation
    \item details of stack destruction
\end{itemize}

\input{continuations}
\input{callcc}
\input{design}
\input{performance}
\input{use_cases}
\input{cc_vs_uc}
\input{implementation}
\input{api}
\input{notes}

%//////////////////////////////////////////////////////////////////////////////

\input{references}

%//////////////////////////////////////////////////////////////////////////////

\end{document}
